---
title: "Varroa EAG response to various essential oils"
author: "Nurit Eliash"
date: "3/16/2022"
output:
  html_document:
    code_folding: hide
    theme: cerulean
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: yes
      df_print: paged
---
<style type="text/css">
  body{
  font-size: 12pt;
}
</style>


## Load libraries
```{r libraries, echo=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(ggthemes)
library(stats) 
library(moments)
library(grid)
library(formattable)
library(gridExtra)
library(ggsignif)
library(hrbrthemes)
library(plotrix)
library(rstatix)
library(car)
library(plotly)
library(postHoc)
knitr::opts_chunk$set(echo = TRUE)
```

## Load data
```{r echo=TRUE, message=FALSE, warning=FALSE}
data <- read.csv("/Users/nuriteliash/Documents/GitHub/Sofia-varroa-project/EAG_data_all.csv")
```

## EAG stat analysis
the tested effect (x) is the Essential oil "EO" dissolved in acetone, the "response" (y) is the electrophysiological response of the leg measured in mV. Each leg was stimulated by different stimuli in the following order:Air > Air > Air > Acetone (solvent) > 0.1 > 0.25 > 0.5 > 1 > 2.5 > 5 > Acetone (solvent) > Air

the response was normalized.... [please complete in here, by copy-pasting from the method section:)]

to test the difference in response amplitude to the different stimuli, we will use One way ANOVA. then a post-hoc Tukey test, to see which of the stimuli are significantly differnet.
we followed this tutorial for [ANOVA in R](http://www.sthda.com/english/wiki/two-way-anova-test-in-r)

### Data summary
```{r echo=TRUE, message=FALSE, warning=FALSE}
# our data contains: 
str(data)
table(data$stimuli, data$EO)
```
We have tested a total of 9 essential oils, using 83 legs. Each essential oil was tested on at least 7 different legs.

### Prior to analysis, we gonna check the ANOVA assumptions:
#### (1) Outliers
```{r echo=TRUE, message=FALSE, warning=FALSE}
# plot it to detect outliers by specific leg
# first sort the order of stimuli:
data$stimuli <- factor(data$stimuli,levels = c("acet_before", "0.1", "0.25", "0.5", "1", "2.5", "5","acet_after"))

box <- ggplot(data, aes(x = stimuli, y = response)) +
  geom_boxplot(aes(colour = EO)) +
  facet_wrap( ~ EO) +
  theme_linedraw() +
        ggtitle("Varroa foreleg response to different essential oils") +
        xlab("Stimuli amount (microgram)") +
        ylab("Normalized response (%)") +
        theme(axis.text.x=element_text(angle=45, hjust=1))
  
ggplotly(box, tooltip = "leg")
```

Remove outliered legs: "VF71", "VF1", "VF2", "VF4", "VF77", "VF82", "VF24", "VF26", "VF37", "VF39", "VF46", "VF49", "VF17", "VF19", "VF50", "VF62"
```{r echo=TRUE, message=FALSE, warning=FALSE}
data <- data %>%
  dplyr::filter(!leg %in% c("VF71", "VF1", "VF2", "VF4", "VF77", "VF82", "VF24", "VF26", "VF37", "VF39", "VF46", "VF49", "VF17", "VF19", "VF50", "VF62"))
```

After excluding the outliers, we proceed with the rest of the tests:

#### (2) Normality:
```{r echo=TRUE, message=FALSE, warning=FALSE}
#the dependent variable should be approximately normally distributed in each cell of the design. This can be checked using the Shapiro-Wilk normality test (shapiro_test()
normality <- data %>%
  group_by(stimuli,EO) %>%
  shapiro_test(response)

normality %>%
  dplyr::filter(p<0.05)
```
For the above four EO x stimuli combinations the response dont distribute normally (p<0.05)

#### (3) Homogneity of variance 
```{r echo=TRUE, message=FALSE, warning=FALSE}
# Build the linear model
model <- lapply(split(data, data$EO), function(i){
  lm(response ~ stimuli, data = i)
  })

# now you can Create a QQ plot of residuals of each EO, for example, essential oil EO4309:
ggqqplot(residuals(model$EO4309))

plot(model$EO4309, 1)
```

After checking the 3 assumptions, we decided to go for a non-parametric test, to test the significant difference of each stimuli concentration from the solvent.

we used a non-parametric post hoc test for multiple comparisons.

### Post hoc non-parametric test for all pairwise comparisons, with Benjamini-Hochberg p-value correction
#### Post hoc Kruskal-Wallis rank sum test
```{r echo=TRUE, message=FALSE, warning=FALSE}
KW <- lapply(split(data, data$EO), function(i){
  posthocKW(i$response, i$stimuli, padjust = "BH")
})

KW$EO4309$PvaluesMatrix
KW$EO4444$PvaluesMatrix
KW$EO4891$PvaluesMatrix
KW$EO4899$PvaluesMatrix
KW$EO5068$PvaluesMatrix
KW$EO5663$PvaluesMatrix
KW$EO5714$PvaluesMatrix
KW$EO5749$PvaluesMatrix
KW$EO5763$PvaluesMatrix
```

#### Post hoc Wilcoxon test
```{r echo=TRUE, message=FALSE, warning=FALSE}
Wilcoxon <- lapply(split(data, data$EO), function(i){
  pairwise.wilcox.test(i$response, g = i$stimuli, p.adjust.method = "BH")
})

Wilcoxon$EO4309$p.value
Wilcoxon$EO4444$p.value
Wilcoxon$EO4891$p.value
Wilcoxon$EO4899$p.value
Wilcoxon$EO5068$p.value
Wilcoxon$EO5663$p.value
Wilcoxon$EO5714$p.value
Wilcoxon$EO5749$p.value
Wilcoxon$EO5763$p.value
```

try Post hoc Wilcoxon test, for specific comparisons, and correct with BH
```{r echo=TRUE, message=FALSE, warning=FALSE}
Wilcoxon <- lapply(split(data, data$EO), function(i){
  pairwise.wilcox.test(i$response, g = i$stimuli, p.adjust.method = "BH")
})


#for each EO, compare each dose, to the avg of "acet_before","acet_after", then correct using BH for each EO. so the correction is only for these 6 comparisons, per EO 

EOs <- unique(data$EO) %>% as.character()
list_df <- list()

for (j in EOs) {
# choose one EO
EO <- data %>% filter(EO == j) 

# calculate the avg response to solvent per EO
before <- EO  %>% filter(stimuli == "acet_before") 
after <- EO  %>% filter(stimuli == "acet_after") 
sol<-left_join(before, after, by="leg") 
sol$response = rowMeans(sol[,c("response.x","response.y")] )

# add the mean response to solvent
df <- sol %>% dplyr::select(c("leg","response","EO"="EO.x")) %>% mutate("stimuli"="solvent")

list_df[[j]] = df
}


#make a loop for Wilcox test of avg solvent vs each of the doses

# state the variables
dose = c("0.1","0.25","0.5","1","2.5","5")
pvalues = vector()
p_list=list()

EOs <- names(list_df)

for (i in dose) {
  for (j in EOs) {
  EO_df <- data %>% filter(EO == j)  %>% filter(stimuli == i) 
  EO_df <- bind_rows(EO_df,list_df[[j]])

  # calculate the p-val for the difference between the response to solvent and the response to stimuli
  pval <-  wilcox.test(data=EO_df , response ~ stimuli, paired = FALSE)$p.value
  pvalues[i] = pval
  
  
  }
  # adjust the pvalues per EO
  padj <- p.adjust(pvalues, "BH")
  wilcox <- data_frame("dose" = dose, "p.value" = pvalues, "padj" = padj, "EO" = EO_df[1,3])
  p_list[[j]] <- wilcox
}



p_list$EO4444
p_list$EO5714
class(list_df$"EO4444")
# somthing is not working, it doesnt run throu all doeses
```

## EAG plot
for some reason, i cannot add the leg ID to the hovering text in the box plot, but i can do it in the dot-plot.
so in order to detect outliers, we can detect them in the first plot (the box plot), then identify their ID leg in the dot plot:)
```{r echo=TRUE, message=FALSE, warning=FALSE}
dot <- ggplot(data, aes(x = stimuli, y = response, colour = EO, leg=leg)) +
  geom_point() + 
  facet_wrap( ~ EO) +
  theme_linedraw() +
        ggtitle("Varroa foreleg response to different essential oils") +
        xlab("Stimuli amount (microgram)") +
        ylab("Normalized response (%)") +
        theme(axis.text.x=element_text(angle=45, hjust=1)) +
        stat_summary(aes(group=EO), fun=mean, geom="line", colour="black") 

ggplotly(dot, tooltip = c("leg","response")) 
```
